Statistical Searches for Microlensing Events in Large, Non-Uniformly Sampled Time-Domain Surveys: A Test Using Palomar Transient Factory Data

### 2013-10-07 text/paper/ms.tex
fixed fig 5 caption
updated some minor grammar

As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}). The PTF survey footprint is not uniformly sampled either spatially or temporally. Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler. The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}. The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry. Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period. The PTF data is not publicly available, but ... TODO

@@

\begin{figure}[h]
\centering\includegraphics[width=.48\textwidth, trim=130 85 90 65, clip]{fig5.eps}
\caption{The maximally outlying light curves for each variability index applied to data from PTF field 100101. For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field. For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field. } \label{fig:indices_examples}
\end{figure}

@@

We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve. In detail, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$. Out tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip --- because of this, we neglect including an AIC or BIC correction to make this a proper likelihood test as the number of data points and number of model parameters are constant. For the Gaussian fit, the optimizer is initialized with a standard-deviation of 10 days, centered on the brightest data point. This is the standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.

>>>>

As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}). The PTF survey footprint is not uniformly sampled either spatially or temporally. Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler. The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}. The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry. Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period. The PTF data are not publicly available, but ... TODO

@@

\begin{figure}[h]
\centering\includegraphics[width=.48\textwidth, trim=10 50 50 50, clip]{fig5.eps}
\caption{Light curves selected from PTF field 100101 with maximally outlying values for each variability index. For $\eta$ and $K$, this corresponds to the light curve with the \emph{minimum} value of the index over the entire field. For the other indices, these are the light curves with the \emph{maximum} value of the relevant index over the field. These light curves represent the type of variability that each index is most sensitive to. } \label{fig:indices_examples}
\end{figure}

@@

We add one more index, $\Delta \chi^2$, the difference in $\chi^2$ between fitting a Gaussian model and fitting a linear model to a light curve. In detail, we use a Levenberg-Marquardt optimizer to perform a least-squares fit with each of these models to the light curves and then compute $\Delta\chi^2=\chi_{linear} - \chi_{gaussian}$. Out tests with $\Delta \chi^2$ below compare the distribution of values over light curves on the same chip --- because of this, we neglect including an AIC or BIC correction to make this a proper likelihood test as the number of data points and number of model parameters are constant. For the Gaussian fit, the optimizer is initialized with a standard-deviation of 10 days, centered on the brightest data point. This is a standard statistical test used by microlensing surveys, and allows us to compare the relative performance of the (slightly modified) \cite{shin2009} indices and of this approach.
